---
title: "3 ToxVal Data Compilation"
author: "Katie Paul Friedman"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    collapsed: yes
    df_print: paged
    lightbox: no
    number_sections: yes
    self_contained: yes
    thumbnails: no
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r setup, warning=FALSE, message=FALSE, echo=FALSE}
library(caret)
library(cowplot)
library(data.table)
library(DescTools)
library(dplyr)
library(DT)
library(ggplot2)
library(ggrepel)
library(ggstance)
library(gplots)
library(httk)
library(kableExtra)
library(jtools)
library(openxlsx)
library(plotly)
library(randomForest)
library(RMySQL)
library(tidyr)
library(tcpl)
library(viridis)
```

```{r, echo=FALSE, eval=FALSE}

con <- dbConnect(drv = RMySQL::MySQL(), user="_dataminer", #fill user
                 password = "pass", # fill pass
                 host = "ccte-mysql-res.epa.gov", 
                 #database = dev_toxval_v9_1)
                 database= res_toxval_v94)
```

# Load Data {.tabset .tabset-fade .tabset-pills}

```{r load-apcra-list, warning=FALSE}

apcra.pro <- fread('./source/chem/apcra_pro.csv') # APCRA prospective chemicals
apcra.pro[,c(1) := NULL]
setnames(apcra.pro, c('V2','V3','V4'), c('DTXSID','CASRN','preferred_name'))
apcra.pro <- apcra.pro[-c(1),]
apcra.pro[,list := 'Pro']

# annotate chemicals that are also in APCRA retrospective case study - these will likely be data-rich
apcra.ret <- as.data.table(read.xlsx('./source/chem/Supp_File_2_pod_ratio_master_final.xlsx', sheet=1)) 
apcra.pro[,apcra.ret := 0]
apcra.ret.dtxsids <- apcra.ret[,DTXSID]
apcra.pro[DTXSID %in% apcra.ret.dtxsids, apcra.ret := 1]

```

```{r load-toxval-94, warning=FALSE, eval=FALSE}
toxval_v9_4 <- dbGetQuery(con, "SELECT * FROM res_toxval_v94.toxval INNER JOIN
                          res_toxval_v94.chemical ON toxval.dtxsid=chemical.dtxsid INNER JOIN
                          res_toxval_v94.species ON toxval.species_id=species.species_id;") %>% as.data.table()

toxval_v9_4 <- toxval_v9_4[dtxsid %in% apcra.pro[,DTXSID]]
save(toxval_v9_4, file='./source/toxval_v9_4_apcra_pro_full.RData')
```

# Refine the toxval data {.tabset .tabset-fade .tabset-pills}

## Human Health

* Select qc_status=='pass'
* Reduce to studies for human health.

```{r, warning=FALSE}
load(file='./source/toxval_v9_4_apcra_pro_full.RData')
```

```{r qc-status-hhra, warning=FALSE}
table(toxval_v9_4$qc_status)
#toxval.apcra <- toxval_v9_1[qa_status==1]
toxval.apcra <- toxval_v9_4[qc_status=='pass']

colnames(toxval.apcra)
head(toxval.apcra)
#table(toxval.apcra$species_supercategory)
table(toxval.apcra$human_eco)
toxval.apcra <- toxval.apcra[human_eco=='human health']

```

```{r select-ra-class, warning=FALSE}

table(toxval.apcra$risk_assessment_class)
toxval.apcra <- toxval.apcra[risk_assessment_class %in% c('chronic',
                                                          'developmental',
                                                          'Hershberger',
                                                          'immunotoxicity',
                                                          'neurotoxicity',
                                                          'neurotoxicity short-term',
                                                          'neurotoxicity subchronic',
                                                          'repeat dose other',
                                                          'reproduction',
                                                          'reproduction developmental',
                                                          'short-term',
                                                          'subchronic',
                                                          'uterotrophic')]

```

```{r select-species, warning=FALSE}

table(toxval.apcra$common_name)

```
* After some inspection it seemed ECOTOX and DOE Wildlife Benchmarks were adding a lot of additional terrestrial species from ecotoxicology.
* However removing these sources would also remove studies from 'Rat' so these seemingly should not be omitted

```{r, warning=FALSE}

toxval.apcra[source %in% c('DOE Wildlife Benchmarks',
                           'ECOTOX')]


```
* Unique listing of species suggests ecotox terrestrial species. 

```{r, warning=FALSE}


unique(toxval.apcra$common_name)

```
* Based on explicitly specifying species, this is what we are left with in the dataset.

```{r specify-species, warning=FALSE}
toxval.apcra <- toxval.apcra[common_name %in% c('Human (RA)',
                                                'Human',
                                                'Rat',
                                                'Mouse',
                                                'Rabbit',
                                                'Dog',
                                                'Rabbit, Rat',
                                                'European Rabbit',
                                                'Mouse, Rat',
                                                'Black Rat',
                                                'Monkey',
                                                'Hamster',
                                                'Guinea Pig',
                                                'Dog, Rat')]

table(toxval.apcra$common_name)
```
* Curious how the human data come to be
* Looks like these are generally specific PPRTV or RfD values
* Can divide these out later as these are probably not fit for comparison to PODs as they are pre-adjusted values (for uncertainty)

```{r what-human, warning=FALSE}

toxval.apcra[common_name %in% c('Human','Human (RA)')]

```


## Repeat exposures 

* How do we want to refine? 
* This case study is about repeat exposures, so drop acute and select repeat dose studies.

```{r, warning=FALSE}

table(toxval.apcra$study_type)
#unique(toxval.apcra$study_type)

toxval.apcra <- toxval.apcra[study_type %in% c('chronic',
                                               'short-term',
                                               'subchronic',
                                               'noncancer',
                                               'developmental',
                                               'repeat dose other',
                                               'reproduction',
                                               'neurotoxicity',
                                               'immunotoxicity',
                                               'uterotrophic',
                                               'neurotoxicity short-term',
                                               'Hershberger',
                                               'reproduction developmental',
                                               'neurotoxicity subchronic'
                                              )]
unique(toxval.apcra$study_type)
```

## Units

* Make sure the toxval_type and units are interpretable for our case study.
* Include oral exposures
* Keep the reference dose (RfD) for now, out of interest (can be dropped later for interpretation).

### Exposure Route

```{r, warning=FALSE}

table(toxval.apcra[,c('exposure_method','exposure_route')])

```
* Sometimes the "-" exposure route is % in diet

```{r, warning=FALSE}

table(toxval.apcra[exposure_route=='-', c('exposure_route','exposure_method_original')])


```
* Included exposure routes that were oral or '-' - the units appeared generally interpretable a oral even when route was '-'

```{r exposure-method-filter, warning=FALSE}

toxval.apcra <- toxval.apcra[exposure_route %in% c('-',
                                                   'oral')]

toxval.apcra[exposure_route=='-' & exposure_method %in% c('-')]
```

### Dose

* Examine the units on dose.
```{r, warning=FALSE}

table(toxval.apcra$toxval_units)

```

* First pass filter on dose units

```{r, warning=FALSE}

toxval.apcra <- toxval.apcra[toxval_units %in% c('%',
                                                 '% diet',
                                                 '(mg/kg-day)-1',
                                                 'g/ kg bdwt/d',
                                                 'g/L',
                                                 'mg/kg',
                                                 'mg/kg-day',
                                                 'mg/kg diet',
                                                 'mg/kg diet/d',
                                                 'ppb diet',
                                                 'ppm',
                                                 'ug/kg bdwt',
                                                 'ug/kg bdwt/d')]

table(toxval.apcra$toxval_units)

```
* Convert all preserved units on dose to mg/kg-bw/day.
* species with ppm

```{r, warning=FALSE}

toxval.apcra[,toxval.apcra.unit := toxval_units]
toxval.apcra[,toxval.apcra.mkd := toxval_numeric]
table(toxval.apcra[toxval_units=='ppm']$common_name)
```
* Ensure all toxval_numeric are numeric by examining range

```{r, warning=FALSE}


toxval.apcra <- toxval.apcra[!(toxval_numeric==-999)] # didn't change after filtering for qa_status
range(toxval.apcra$toxval_numeric)

```

```{r, warning=FALSE}

table(toxval.apcra[,c('toxval_units')])
```

* Convert all units to mg/kg/day
###

```{r ug-kg-mkd, warning=FALSE}
# conversion from ug or g/kg-day to mg/kg-day (mkd)
toxval.apcra[toxval_units=='ug/kg bdwt', toxval.apcra.mkd := toxval_numeric/1000]
toxval.apcra[toxval_units=='ug/kg bdwt/d', toxval.apcra.mkd := toxval_numeric/1000]
toxval.apcra[toxval_units %in% c('ug/kg bdwt','ug/kg bdwt/d'), toxval.apcra.unit := 'mg/kg-day']
```

* Based on subset of % and % diet, will assume that is feed and delete Hamster

```{r, warning=FALSE}

table(toxval.apcra[toxval_units %in% c('% diet','%'), c('exposure_method','common_name')])
# will assume that % means in feed 
```
```{r pct-diet-mkd, warning=FALSE}

# conversions for % feed/diet into mkd, where 1% in diet = 10,000 ppm
# get from % in diet to ppm multiplying by 10,000
# get to mg/kg/d by muliplying by the 1ppm to mkd conversion by species

toxval.apcra[toxval_units %in% c('%', '% diet') & common_name %in% c('dog','Dog'), toxval.apcra.mkd := (toxval_numeric*10000)*0.025]
toxval.apcra[toxval_units %in% c('%', '% diet') & common_name %in% c('rat', 'Rat'), toxval.apcra.mkd := (toxval_numeric*10000)*0.05]
toxval.apcra[toxval_units %in% c('%', '% diet') & common_name %in% c('mouse', 'Mouse'), toxval.apcra.mkd := (toxval_numeric*10000)*0.15]
toxval.apcra[toxval_units %in% c('%', '% diet') & common_name %in% c('rabbit','Rabbit'), toxval.apcra.mkd := (toxval_numeric*10000)*0.03]
toxval.apcra <- toxval.apcra[!(toxval_units %in% c('%', '% diet') & common_name %in% c('Hamster'))]

toxval.apcra[toxval_units %in% c('%', '% diet'), toxval.apcra.unit := 'mg/kg-day']

```

```{r, warning=FALSE}

table(toxval.apcra[toxval_units %in% c('ppm','ppb'),c('toxval_units','common_name')])
```


```{r convert-ppm-mkd, warning=FALSE}
# conversions to mkd from ppm

#toxval.apcra[toxval_units=='ppm' & species_common=='dog', toxval.apcra.mkd := toxval_numeric*0.025]
toxval.apcra[toxval_units=='ppm' & common_name=='Rat', toxval.apcra.mkd := toxval_numeric*0.05]
toxval.apcra[toxval_units=='ppm' & common_name=='Black Rat', toxval.apcra.mkd := toxval_numeric*0.05]
toxval.apcra[toxval_units=='ppm' & common_name=='Mouse', toxval.apcra.mkd := toxval_numeric*0.15]
#toxval.apcra[toxval_units=='ppm' & species_common=='rabbit', toxval.apcra.mkd := toxval_numeric*0.03]
#toxval.apcra[toxval_units=='ppm' & species_common=='guinea pig', toxval.apcra.mkd := toxval_numeric*0.040]
toxval.apcra[toxval_units=='ppm' & common_name=='Hamster', toxval.apcra.mkd := toxval_numeric*0.094]
toxval.apcra[toxval_units=='ppm', toxval.apcra.unit := 'mg/kg-day']
```

```{r, warning=FALSE}
# conversions from ppb diet
table(toxval.apcra[toxval_units=='ppb diet', c('common_name', 'toxval_units')])

```
```{r convert ppb-mkd, warning=FALSE}

toxval.apcra[toxval_units=='ppb diet' & common_name=='Mouse',toxval.apcra.mkd := toxval_numeric*0.001*0.015 ]
toxval.apcra[toxval_units=='ppb diet' & common_name=='Rat',toxval.apcra.mkd := toxval_numeric*0.001*0.05 ]
toxval.apcra[toxval_units=='ppb diet', toxval.apcra.unit := 'mg/kg-day']
```

```{r, warning=FALSE}

table(toxval.apcra[toxval_units %in% c('mg/kg diet','mg/kg diet/d'),c('toxval_units','common_name')])

# toxval.apcra[toxval_units=='mg/kg'] # somewhat impossible to know if this is in diet or kg-bw; on inspection it mostly seems like mg/kg-bw
```

```{r convert-mg-kg-diet-mkd,warning=FALSE}

# conversions for mg/kg diet to mkd, where 1 mg/kg diet ~ 1 ppm in diet
toxval.apcra[toxval_units=='mg/kg diet' & common_name=='Dog', toxval.apcra.mkd := toxval_numeric*0.025]
toxval.apcra[toxval_units=='mg/kg diet' & common_name=='Rat', toxval.apcra.mkd := toxval_numeric*0.05]
toxval.apcra[toxval_units=='mg/kg diet/d' & common_name=='Rat', toxval.apcra.mkd := toxval_numeric*0.05]
toxval.apcra[toxval_units=='mg/kg diet' & common_name=='Mouse', toxval.apcra.mkd := toxval_numeric*0.15]
toxval.apcra[toxval_units=='mg/kg diet' & common_name=='Hamster', toxval.apcra.mkd := toxval_numeric*0.094]
toxval.apcra[toxval_units %in% c('mg/kg diet','mg/kg diet/d'), toxval.apcra.unit := 'mg/kg-day']

```


```{r, warning=FALSE}

toxval.apcra[toxval_units=='g/L' & common_name=='Rat', toxval.apcra.mkd := toxval_numeric*1000*0.05]
toxval.apcra[toxval_units %in% c('g/L'), toxval.apcra.unit := 'mg/kg-day']
```

```{r, warning=FALSE}

table(toxval.apcra$toxval.apcra.unit)
```

### ToxVal type

```{r, warning=FALSE}
table(toxval.apcra$toxval_type)
#unique(toxval.apcra$toxval_type)
```

* Keep only the following units on effect level

```{r, warning=FALSE}

toxval.apcra <- toxval.apcra[toxval_type %in% c('BMD',
                                                'BMDL',
                                                'BMDL10',
                                                'HNEL',
                                                'LEL',
                                                'LOAEL',
                                                'NEL',
                                                'NOAEL',
                                                'NOAEL ',
                                                'NOEL',
                                                'NEL',
                                                # remove these later but want to see them first
                                                'RfD',
                                                'cancer slope factor'
                                                )]
```

```{r, warning=FALSE}

table(toxval.apcra$toxval_type)

```

## Examine doses

* A histogram of the values suggests that most fall within 1-1000 mg/kg-bw/day. 
* likely want to examine extreme values.
* Most of the extreme values appear to be RfDs (corrected with UFs).

```{r, warning=FALSE}


ggplot(data=toxval.apcra[,c('toxval.apcra.mkd')], aes(x=log10(toxval.apcra.mkd)))+
  geom_histogram(bins=50)+
  theme_bw()+
  xlab("Toxval log10-mg/kg/day values")
  

```

* Remove RfDs and look at distribution of NELs/NOAELs/LOAELs/LELs

```{r, warning=FALSE}

toxval.apcra.norfds <- toxval.apcra[!(toxval_type %in% c('RfD','cancer slope factor'))]
ggplot(data=toxval.apcra.norfds[,c('toxval.apcra.mkd')], aes(x=log10(toxval.apcra.mkd)))+
  geom_histogram(bins=50)+
  theme_bw()
  

```

* 262 observations are less than 0.1 mg/kg/day
* 51 observations are less than 0.01 mg/kg/day for 9 dtxsids
* 165 DTXSIDs in the dataset with no RfDs

```{r, warning=FALSE}

lo.values <- toxval.apcra.norfds[toxval.apcra.mkd < 0.01]

length(unique(toxval.apcra.norfds$dtxsid)) #165 dtxsids
```

* One chemical in the low observations appears to have one observation that is much lower than the others for the chemical.

```{r, warning=FALSE}
lo.values[dtxsid=='DTXSID6020062']

```

* Look at distribution just for this one chemical...DTXSID6020062
* the lowest values are all the same, seem to come from same study, and are extremely low compared to the rest.

```{r, warning=FALSE}

ggplot(data=toxval.apcra.norfds[dtxsid=='DTXSID6020062',c('toxval.apcra.mkd')], aes(x=log10(toxval.apcra.mkd)))+
  geom_histogram(bins=50)+
  theme_bw()

```

# Compare the new ToxVal PODs {.tabset .tabset-fade .tabset-pills}

## Create summary

* Create the summary for comparison to the previous retrospective chemicals.
* Retrospective case study used a refined set from dev_toxval_v5.
* Adding in summary quantile values from 5-30%.

```{r, warning=FALSE}
toxval.apcra.summary <- unique(toxval.apcra.norfds[, list(
  min.toxval.numeric = min(toxval.apcra.mkd),
  p5.toxval.numeric = quantile(toxval.apcra.mkd, probs=c(0.05)),
  p10.toxval.numeric = quantile(toxval.apcra.mkd, probs=c(0.10)),
  p15.toxval.numeric = quantile(toxval.apcra.mkd, probs=c(0.15)),
  p20.toxval.numeric = quantile(toxval.apcra.mkd, probs=c(0.20)),
  p25.toxval.numeric = quantile(toxval.apcra.mkd, probs=c(0.25)),
  p30.toxval.numeric = quantile(toxval.apcra.mkd, probs=c(0.30)),
  max.toxval.numeric = max(toxval.apcra.mkd),
  median.toxval.numeric = median(toxval.apcra.mkd),
  mean.toxval.numeric = mean(toxval.apcra.mkd),
  stdev.toxval.numeric = sd(toxval.apcra.mkd),
  number.toxval.numeric = .N
), by=list(dtxsid, casrn, name)])

```

* Create separate summary for subchronic only.

```{r, warning=FALSE}

table(toxval.apcra.norfds$risk_assessment_class)
```

```{r, warning=FALSE, echo=FALSE, eval=FALSE}
# discovered a bug in the code where we were using study_duration_class instead of risk_assessment_class to pull subchronic
# can see there are a lot of subchronics from risk_assessment_class and not many from study_duration_class
table(toxval.apcra.norfds$study_duration_class)

```

```{r, warning=FALSE}
# updating from study_duration_class to risk_assessment_class increases the number of chemicals with subchronic PODs from 9 to 160

toxval.apcra.summary.subchronic <- unique(toxval.apcra.norfds[risk_assessment_class %in% c('subchronic','repeat dose other','short-term'), list(
  min.toxval.numeric.sub = min(toxval.apcra.mkd),
  p5.toxval.numeric.sub = quantile(toxval.apcra.mkd, probs=c(0.05)),
  p10.toxval.numeric.sub = quantile(toxval.apcra.mkd, probs=c(0.10)),
  p15.toxval.numeric.sub = quantile(toxval.apcra.mkd, probs=c(0.15)),
  p20.toxval.numeric.sub = quantile(toxval.apcra.mkd, probs=c(0.20)),
  p25.toxval.numeric.sub = quantile(toxval.apcra.mkd, probs=c(0.25)),
  p30.toxval.numeric.sub = quantile(toxval.apcra.mkd, probs=c(0.30)),
  max.toxval.numeric.sub = max(toxval.apcra.mkd),
  median.toxval.numeric.sub = median(toxval.apcra.mkd),
  mean.toxval.numeric.sub = mean(toxval.apcra.mkd),
  stdev.toxval.numeric.sub = sd(toxval.apcra.mkd),
  number.toxval.numeric.sub = .N
), by=list(dtxsid, casrn, name)])

```

* APCRA retrospective supp file 2 values
* Use the 5th percentile PODs for comparison
* 96 substances to compare

```{r, warning=FALSE}

toxval.apcra.summary$apcra.ret.5p.POD <- apcra.ret$p5.POD[match(toxval.apcra.summary$dtxsid,
                                                                apcra.ret$DTXSID)]

toxval.apcra.summary[, diff := as.numeric(log10(p5.toxval.numeric) - log10(apcra.ret.5p.POD)) ]
toxval.apcra.summary[!is.na(diff),label := name]
toxval.apcra.summary[diff > -0.5, label := ""]
toxval.apcra.summary[is.na(diff), label := ""]

big.diffs <- toxval.apcra.summary[diff< -0.5, dtxsid]

length(unique(toxval.apcra.summary[!is.na(diff)]$dtxsid))

```

## Comparison of overall POD to retrospective case study POD

```{r, warning=FALSE, fig.height=4, fig.width=6}
#library(ggrepel)

ggplot(data=toxval.apcra.summary, aes(x=apcra.ret.5p.POD, y=p5.toxval.numeric)) +
  geom_point(size=1)+ 
  geom_text_repel(data=toxval.apcra.summary, aes(x=apcra.ret.5p.POD, y=p5.toxval.numeric, label=label))+
  scale_y_log10(limits=c(10^-3,10^5),
                breaks = scales::trans_breaks("log10", function(x) 10^x),
                labels = scales::trans_format("log10", scales::math_format(10^.x)))+
   scale_x_log10(limits=c(10^-3,10^5),
                breaks = scales::trans_breaks("log10", function(x) 10^x),
                labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  geom_abline(slope=1, intercept=0) +
  geom_abline(slope=1, intercept=0.5, linetype="dashed") +
  geom_abline(slope=1, intercept=-0.5, linetype="dashed")+
  theme_bw()+
  theme(axis.title = element_text(size=14),
        axis.text = element_text(size=12))+
  xlab('5th percentile APCRA retrospective POD')+
  ylab('5th percentile APCRA prospective POD')


```
* For these substances, the new ToxVal 5th percentile POD was lower by at least 0.5 log10-mg/kg/day
* None of them seem quite as egregious as the one above - no separations of 3-4 orders of magnitude between lowest values and next lowest.

```{r, warning=FALSE, fig.height=6, fig.width=8}

diffs <- toxval.apcra.norfds[dtxsid %in% big.diffs, c('dtxsid','name','toxval.apcra.mkd')]
ggplot(data=diffs, aes(x=log10(toxval.apcra.mkd)))+
  geom_histogram(bins=50)+
  theme_bw()+
  scale_y_log10()+
  facet_wrap(~ name)
```

## Comparison of different quantiles

```{r, warning=FALSE}

toxval.summary.long <- melt.data.table(toxval.apcra.summary,
                                       id.vars = c('dtxsid','casrn','name', 'apcra.ret.5p.POD'),
                                       measure.vars = c('p5.toxval.numeric',
                                                        'p10.toxval.numeric',
                                                        'p15.toxval.numeric',
                                                        'p20.toxval.numeric',
                                                        'p25.toxval.numeric',
                                                        'p30.toxval.numeric'),
                                       variable.name = c('toxval.p'))

col.num <- c("value", "apcra.ret.5p.POD")
toxval.summary.long <- toxval.summary.long[, (col.num) := lapply(.SD, function(x) log10(x)), .SDcols = col.num ]
head(toxval.summary.long)
```


```{r fig.width=10, fig.height=6, warning=FALSE}

fig.toxval.ecdf <- ggplot(toxval.summary.long[toxval.p %in% c('p5.toxval.numeric',
                                                        'p10.toxval.numeric',
                                                        'p15.toxval.numeric',
                                                        'p20.toxval.numeric',
                                                        'p25.toxval.numeric',
                                                        'p30.toxval.numeric')], aes(value, color=toxval.p))+
  stat_ecdf(geom='step', size=1.5)+
  scale_y_continuous(trans = 'log10',
                     breaks= c(0.01, 0.1,0.2,0.3,0.4,0.5,0.75, 1))+
  ylab("Cumulative Frequency") +
  xlab('log10 POD quantile')+
  theme_bw() +
  theme(
    axis.line = element_line(colour = "black"),
    axis.text = element_text(size=10),
    axis.title = element_text(size=12, face='bold'))+
  theme(axis.text.y = element_text(family = "sans", face = "bold", size=12))+
  theme(legend.position="right", legend.title=element_blank())+
  scale_x_continuous(breaks=seq(-5,10,1)) +
  coord_cartesian(xlim = c(-5, 10)) +
  #scale_color_viridis(discrete=TRUE, name='Ratio Type')+
  scale_colour_manual(breaks=c('p5.toxval.numeric',
                                                        'p10.toxval.numeric',
                                                        'p15.toxval.numeric',
                                                        'p20.toxval.numeric',
                                                        'p25.toxval.numeric',
                                                        'p30.toxval.numeric'),
                      values=c("#440154FF", "#3B528BFF", "#21908CFF", "#5DC863FF", "#FDE725FF", 'darkmagenta'),
                      labels('5th %-ile','10th %-ile','15th %-ile', '20th %-ile','25th %ile', '30th %ile'))+
  geom_vline(xintercept=-2, lty='dashed', color='red')+
  geom_vline(xintercept=2, lty='dashed', color='red')+
  geom_vline(xintercept=0, color='red')+
  geom_hline(yintercept=0.90, lty='dashed', color='red')

fig.toxval.ecdf




```



```{r, warning=FALSE}

fig.toxval.violin <- ggplot(toxval.summary.long[toxval.p %in% c('p5.toxval.numeric',
                                                        'p10.toxval.numeric',
                                                        'p15.toxval.numeric',
                                                        'p20.toxval.numeric',
                                                        'p25.toxval.numeric',
                                                        'p30.toxval.numeric')], aes(x=toxval.p, y=value, fill=toxval.p))+
  geom_violin(draw_quantiles = c(0.25,0.5,0.75), color='white', trim=FALSE)+
  #geom_boxplot(width=0.1, color='white')+
  scale_y_continuous(breaks=seq(-3,2,0.5))+
  ylab("ToxVal POD Value, log10-mg/kg/day") +
  xlab('Percentile')+
  theme_bw() +
  theme(
    axis.line = element_line(colour = "black"),
    axis.text = element_text(size=10),
    axis.title = element_text(size=12, face='bold'))+
  theme(
        axis.text.x = element_text(angle=45, hjust=1))+
  theme(legend.position="right", legend.title=element_blank())+
  scale_fill_manual(values=viridis(6),
                    breaks=c('p5.toxval.numeric',
                                                        'p10.toxval.numeric',
                                                        'p15.toxval.numeric',
                                                        'p20.toxval.numeric',
                                                        'p25.toxval.numeric',
                                                        'p30.toxval.numeric'),
                    labels=c('5th','10th','15th','20th','25th','30th'))+
  scale_x_discrete(labels=c('p5.toxval.numeric' ='5th',
                                                        'p10.toxval.numeric' = '10th',
                                                        'p15.toxval.numeric' = '15th',
                                                        'p20.toxval.numeric' = '20th',
                                                        'p25.toxval.numeric' = '25th',
                                                        'p30.toxval.numeric' = '30th'))

fig.toxval.violin
```
```{r, warning=FALSE}

wilcox <- group_by(toxval.summary.long, toxval.p) %>%
  summarise(
    count = n(),
    median = median(value, na.rm = TRUE),
    IQR = IQR(value, na.rm = TRUE)
  )
wilcox
```


```{r, warning=FALSE}

wilcox.test(data=toxval.apcra.summary, 
            toxval.apcra.summary$p5.toxval.numeric, 
            toxval.apcra.summary$p10.toxval.numeric, mu=0, alt='two.sided',paired=TRUE, conf.int=TRUE, conf.level=0.95)
```

```{r, warning=FALSE}

fig.toxval.distrib <- ggplot()+
  geom_histogram(data=toxval.apcra.norfds[,c('dtxsid','toxval.apcra.mkd')],
                 aes(x=toxval.apcra.mkd))+
  
  theme_bw()+
  ylab('Frequency')+
  xlab('ToxVal Numeric POD (log10-mg/kg/day)')+
  theme(axis.text.x = element_text(size=12, angle=45, hjust=1),
        axis.text.y = element_text(size=12),
        axis.title = element_text(size=14),
        plot.title = element_text(hjust=0.5))+
  scale_x_log10(limits=c(10^-4,10^4),
                breaks = scales::trans_breaks("log10", function(x) 10^x),
                labels = scales::trans_format("log10", scales::math_format(10^.x)))


fig.toxval.distrib
```



```{r, warning=FALSE, eval=FALSE}

file.dir <- paste('output/', sep='')
file.name <- paste('/SuppFig_ToxVal_percentiles_', Sys.Date(), '.tiff', sep='')
file.path <- paste(file.dir, file.name, sep='')
dir.create(path=file.dir, showWarnings = FALSE, recursive = TRUE)
tiff(file.path, width=8, height=5, units='in', res=450)
plot_grid(fig.toxval.distrib, fig.toxval.violin, ncol=2, labels=c("A", "B"), label_size = 14)
dev.off()


```

# Reproducibility {.tabset .tabset-fade .tabset-pills}

## Write files

```{r, warning=FALSE, eval=FALSE}


list_data <- list("toxval.apcra.pro.summary" = as.data.frame(toxval.apcra.summary),
                  "toxval.apcra.pro.summ.SUBonly" = as.data.frame(toxval.apcra.summary.subchronic),
                  "toxval.apcra.pro.all" = as.data.frame(toxval.apcra.norfds))

write.xlsx(list_data, './source/apcra_pro_toxval_v9_4_PODs.xlsx')
save(toxval.apcra.norfds,
     toxval.apcra.summary,
     toxval.apcra.summary.subchronic,
     file='./source/apcra_pro_toxval_v9_4_PODs.RData')

```

## Rsession

```{r, warning=FALSE}

print(sessionInfo())

```